openTSNE: a modular \proglang{Python} library for t-SNE dimensionality reduction and embedding

### 2022-03-09 paper/main.tex
Update package versions

Fig.~\ref{fig:benchmarks} benchmarks four popular Python t-SNE implementations,
including those from \pkg{scikit-learn} (v0.23.1), \pkg{MulticoreTSNE} (v0.1),
\pkg{FIt-SNE} (v1.1.0), and our \opentsne\ (v0.6.0). We perform benchmarks on
two computational platforms, one representing the usage on a personal computer
and the other the utility of these libraries on a high-performance computing
platform. The Intel(R) Core i7 is commonly found in consumer-grade laptop
computers, while Intel(R) Xeon(R) processors appear in high-performance
computing machines. Benchmarks were run for $1,000$ iterations with the original
t-SNE parameters, as some implementations do not allow their modification.

\begin{figure*}[ht]
  \includegraphics[width=\textwidth]{benchmarks}
  \caption{\label{fig:benchmarks}
  We benchmark \opentsne\ (v0.6.0) against three popular open-source
  implementations from \pkg{scikit-learn}~\citep{pedregosa2011scikit}
  (v0.23.1), \pkg{MulticoreTSNE}~\citep{Ulyanov2016} (v0.1), and
  \pkg{FIt-SNE}~\citep{linderman2019fast} (v1.1.0). Experiments were run on a
  consumer-grade Intel Core i7-7700HQ processor found in laptop computers, and
  on a server-grade Intel Xeon E5-2650. To generate benchmark data sets of
  different sizes, we subsampled data from the 10X Genomics 1.3 million mouse
  brain data set five times, resulting in five different data sets for each
  size. In total, we run each implementation on 30 different data sets. Notice
  that \opentsne\ scales similarly to \pkg{FIt-SNE}, as they both use the
  same interpolation-based approximation scheme, while \pkg{scikit-learn} and
  \pkg{MulticoreTSNE} utilize the Barnes-Hut approximation.
}
\end{figure*}

@@

\caption{\label{tab:features}
  We compare the features of \opentsne~(v0.6.0) to four popular open-source
  t-SNE implementations from \pkg{scikit-learn}~(v0.23.1),
  \pkg{BHTSNE}~(master), \pkg{MulticoreTSNE}~(v0.1), and
  \pkg{FIt-SNE}~(v1.1.0).
}
\end{table}

@@

The results in this paper were obtained using \proglang{Python}~3.7.6 with the
\pkg{openTSNE}~v0.6.0 package. \opentsne\ has three direct dependencies:
\pkg{numpy}, \pkg{scipy}, and \pkg{scikit-learn}. In our experiments, we used
\pkg{numpy}~v1.18.1, \pkg{scipy}~v1.4.1. \pkg{scikit-learn}~v0.23.1. As these
packages include their own downstream dependencies, the full environment used to
produce the figures is included in the accompanying GitHub repository
\url{https://github.com/pavlin-policar/opentsne-paper}.

>>>>

Fig.~\ref{fig:benchmarks} benchmarks four popular Python t-SNE implementations,
including those from \pkg{scikit-learn} (v1.0.1), \pkg{MulticoreTSNE} (v0.1),
\pkg{FIt-SNE} (v1.1.0), and our \opentsne\ (v0.6.1). We perform benchmarks on
two computational platforms, one representing the usage on a personal computer
and the other the utility of these libraries on a high-performance computing
platform. The Intel(R) Core i7 is commonly found in consumer-grade laptop
computers, while Intel(R) Xeon(R) processors appear in high-performance
computing machines. Benchmarks were run for $1,000$ iterations with the original
t-SNE parameters, as some implementations do not allow their modification.

\begin{figure*}[ht]
  \includegraphics[width=\textwidth]{benchmarks}
  \caption{\label{fig:benchmarks}
  We benchmark \opentsne\ (v0.6.1) against three popular open-source
  implementations from \pkg{scikit-learn}~\citep{pedregosa2011scikit}
  (v1.0.1), \pkg{MulticoreTSNE}~\citep{Ulyanov2016} (v0.1), and
  \pkg{FIt-SNE}~\citep{linderman2019fast} (v1.1.0). Experiments were run on a
  consumer-grade Intel Core i7-7700HQ processor found in laptop computers, and
  on a server-grade Intel Xeon E5-2650. To generate benchmark data sets of
  different sizes, we subsampled data from the 10X Genomics 1.3 million mouse
  brain data set five times, resulting in five different data sets for each
  size. In total, we run each implementation on 30 different data sets. Notice
  that \opentsne\ scales similarly to \pkg{FIt-SNE}, as they both use the
  same interpolation-based approximation scheme, while \pkg{scikit-learn} and
  \pkg{MulticoreTSNE} utilize the Barnes-Hut approximation.
}
\end{figure*}

@@

\caption{\label{tab:features}
  We compare the features of \opentsne~(v0.6.1) to four popular open-source
  t-SNE implementations from \pkg{scikit-learn}~(v1.0.1),
  \pkg{BHTSNE}~(master), \pkg{MulticoreTSNE}~(v0.1), and
  \pkg{FIt-SNE}~(v1.1.0).
}
\end{table}

@@

The results in this paper were obtained using \proglang{Python}~3.9.7 with the
\pkg{openTSNE}~v0.6.1 package. \opentsne\ has three direct dependencies:
\pkg{numpy}, \pkg{scipy}, and \pkg{scikit-learn}. In our experiments, we used
\pkg{numpy}~v1.21.3, \pkg{scipy}~v1.7.1. \pkg{scikit-learn}~v1.0.1. As these
packages include their own downstream dependencies, the full environment used to
produce the figures is included in the accompanying GitHub repository
\url{https://github.com/pavlin-policar/opentsne-paper}.
